---
output: github_document
bibliography: inst/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

> crumble (verb): break or fall apart into small fragments

# crumble

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/crumble)](https://CRAN.R-project.org/package=crumble) [![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

<!-- badges: end -->

*crumble* implements a modern, unified estimation strategy [@liu2024general] for common mediation estimands: natural effects [@pearl2022], organic effects [@lok2015], interventional effects [@vansteelandt2017], recanting twins [@vo2024], in causal inference in combination with modified treatment policies. It makes use of recent advancements in "Riesz-learning" to estimate a set of required nuisance parameters using deep learning. The result is a software package that is capable of estimating mediation effects with binary, categorical, continuous, or multivariate exposures with high-dimensional mediators and mediator-outcome confounders using machine learning.

### Installation

```{r eval=FALSE}
remotes::install_github("nt-williams/crumble")
```

### Features

| Feature                   | Status  |
|---------------------------|:-------:|
| Recanting twins           |    ✓    |
| Natural effects           |    ✓    |
| Organic effects           |    ✓    |
| Interventional effects    |    ✓    |
| Modified treatment Policy |    ✓    |
| Static intervention       |    ✓    |
| Dynamic intervention      |    ✓    |
| Continuous treatment      |    ✓    |
| Binary treatment          |    ✓    |
| Categorical treatment     |    ✓    |
| Multivariate treatment    |    ✓    |
| Missingness in treatment  |         |
| Continuous outcome        |    ✓    |
| Binary outcome            |    ✓    |
| Censored outcome          |    ✓    |
| Survey weights            | Planned |
| Super learner             |    ✓    |
| Clustered data            | Planned |
| Parallel processing       |    ✓    |
| GPU support               |    ✓    |
| Progress bars             |    ✓    |

### Example(s)

```{r eval=FALSE}
library(crumble)
library(mlr3extralearners)

data(weight_behavior, package = "mma")

weight_behavior <- na.omit(weight_behavior)

set.seed(2345)
```

##### Recanting twins

```{r eval=FALSE}
crumble(
	data = weight_behavior,
	trt = "sports", 
	outcome = "bmi",
	covar = c("age", "sex", "tvhours"),
	mediators = c("exercises", "overweigh"),
	moc = "snack",
	d0 = \(data, trt) factor(rep(1, nrow(data)), levels = c("1", "2")), 
	d1 = \(data, trt) factor(rep(2, nrow(data)), levels = c("1", "2")), 
	effect = "RT",
	learners = c("mean", "glm", "earth", "ranger"), 
	nn_module = sequential_module(),
	control = crumble_control(crossfit_folds = 1L, epochs = 20L)
)
#> ✔ Permuting Z-prime variables... 1/1 tasks [2.5s]
#> ✔ Fitting outcome regressions... 1/1 folds [25.6s]             
#> ✔ Computing alpha n density ratios... 1/1 folds [39.7s]        
#> ✔ Computing alpha r density ratios... 1/1 folds [41.6s]        
#> 
#> ══ Results `crumble()` ═════════════════════════════════════════
#> 
#> ── E[Y(d1) - Y(d0)] 
#>       Estimate: 1.0537
#>     Std. error: 0.3009
#>         95% CI: (0.4639, 1.6435)
#> 
#> ── Path: A -> Y 
#>       Estimate: 0.0366
#>     Std. error: 0.1842
#>         95% CI: (-0.3245, 0.3976)
#> 
#> ── Path: A -> Z -> Y 
#>       Estimate: -0.0202
#>     Std. error: 0.0238
#>         95% CI: (-0.0668, 0.0264)
#> 
#> ── Path: A -> Z -> M -> Y 
#>       Estimate: -6e-04
#>     Std. error: 0.0099
#>         95% CI: (-0.02, 0.0189)
#> 
#> ── Path: A -> M -> Y 
#>       Estimate: 1.0506
#>     Std. error: 0.2162
#>         95% CI: (0.627, 1.4743)
#> 
#> ── Intermediate Confounding 
#>       Estimate: -0.0127
#>     Std. error: 0.0261
#>         95% CI: (-0.0638, 0.0384)
```

##### Natural effects

```{r eval=FALSE}
crumble(
	data = weight_behavior,
	trt = "sports", 
	outcome = "bmi",
	covar = c("age", "sex", "tvhours"),
	mediators = c("exercises", "overweigh"),
	d0 = \(data, trt) factor(rep(1, nrow(data)), levels = c("1", "2")), 
	d1 = \(data, trt) factor(rep(2, nrow(data)), levels = c("1", "2")), 
	effect = "N",
	learners = c("mean", "glm", "earth", "ranger"), 
	nn_module = sequential_module(),
	control = crumble_control(crossfit_folds = 1L, epochs = 20L)
)
#> ✔ Fitting outcome regressions... 1/1 folds [10.6s]             
#> ✔ Computing alpha n density ratios... 1/1 folds [53.1s]        
#> 
#> ══ Results `crumble()` ═════════════════════════════════════════
#> 
#> ── E[Y(d1) - Y(d0)] 
#>       Estimate: 1.0289
#>     Std. error: 0.28
#>         95% CI: (0.48, 1.5777)
#> 
#> ── Natural Direct Effect 
#>       Estimate: 0.0165
#>     Std. error: 0.1717
#>         95% CI: (-0.3201, 0.3531)
#> 
#> ── Natural Indirect Effect 
#>       Estimate: 1.0124
#>     Std. error: 0.2178
#>         95% CI: (0.5856, 1.4393)
```

##### Organic effects

```{r eval=FALSE}
crumble(
	data = weight_behavior,
	trt = "sports", 
	outcome = "bmi",
	covar = c("age", "sex", "tvhours"),
	mediators = c("exercises", "overweigh"),
	d0 = \(data, trt) factor(rep(1, nrow(data)), levels = c("1", "2")), 
	d1 = \(data, trt) factor(rep(2, nrow(data)), levels = c("1", "2")), 
	effect = "O",
	learners = c("mean", "glm", "earth", "ranger"), 
	nn_module = sequential_module(),
	control = crumble_control(crossfit_folds = 1L, epochs = 20L)
)
#> ✔ Fitting outcome regressions... 1/1 folds [10.7s]             
#> ✔ Computing alpha n density ratios... 1/1 folds [48.2s]        
#> 
#> ══ Results `crumble()` ═════════════════════════════════════════
#> 
#> ── Organic Direct Effect 
#>       Estimate: 0.011
#>     Std. error: 0.1772
#>         95% CI: (-0.3364, 0.3584)
#> 
#> ── Organic Indirect Effect 
#>       Estimate: 1.0278
#>     Std. error: 0.2231
#>         95% CI: (0.5904, 1.4651)#> 
```

##### Randomized interventional effects

```{r eval=FALSE}
crumble(
	data = weight_behavior,
	trt = "sports", 
	outcome = "bmi",
	covar = c("age", "sex", "tvhours"),
	mediators = c("exercises", "overweigh"),
	moc = "snack",
	d0 = \(data, trt) factor(rep(1, nrow(data)), levels = c("1", "2")), 
	d1 = \(data, trt) factor(rep(2, nrow(data)), levels = c("1", "2")), 
	effect = "RI",
	learners = c("mean", "glm", "earth", "ranger"), 
	nn_module = sequential_module(),
	control = crumble_control(crossfit_folds = 1L, epochs = 20L)
)
#> ✔ Permuting Z-prime variables... 1/1 tasks [2s]
#> ✔ Fitting outcome regressions... 1/1 folds [14.2s]             
#> ✔ Computing alpha r density ratios... 1/1 folds [1m 23.2s]     
#> 
#> ══ Results `crumble()` ═════════════════════════════════════════
#> 
#> ── Randomized Direct Effect 
#>       Estimate: 0.0162
#>     Std. error: 0.1774
#>         95% CI: (-0.3315, 0.364)
#> 
#> ── Randomized Indirect Effect 
#>       Estimate: 1.0304
#>     Std. error: 0.2296
#>         95% CI: (0.5805, 0.4662)
```

#### References
